### 2018-07-26 express2018/formal.tex
HOL traces its roots back to Edinburgh LCF by Robin Milner and others in 1972

Higher Order Logic (or \emph{HOL Logic}) \cite{hollogic} is a variant of
Church’s simple theory of types (STT) \cite{church1940formulation}
\hl{(or \emph{simple-typed $\lambda$-calculus}),
plus a higher order version of Hilbert's choice operator $\varepsilon$,
Axiom of Infinity, and Rank-1 (prenex) polymorphism.}
HOL4 has implemented the original HOL Logic, 
while some other theorem provers in HOL family (e.g. Isabelle/HOL) have
certain extensions.
%  (they made the formal language more powerful,
% but they also bring the possibilities that the entire logic becomes
% inconsistent). 
Indeed the HOL Logic has considerable simpler logical
foundations than most other theorem provers. %, e.g. Coq. 
As a result,
formal theories built in HOL4 is relatively easier convincible and can
be also easily migrated into other theorem provers,
sometimes even automatically \cite{hurd2011opentheory}.

HOL4 is written in Standard ML, a single programming language which
plays three different roles:
%(The situation is quite different in other systems like Coq\footnote{\hl{The Coq Proof Assistant.} \url{https://coq.inria.fr}}.)
\begin{enumerate}
\item It serves as the underlying implementation language for the core HOL engine;
\item it is used to implement tactics (and tacticals) for writing proofs;
\item it is used as the command language of the HOL interactive shell.
\end{enumerate}
\hl{Moreover, using} the same language HOL4 users can write complex automatic
verification tools by calling HOL's theorem proving
facilities. \hl{(The formal proofs of theorems in CCS theory
are mostly done by an \emph{interactive process} closely following
their informal proofs, with minimal automatic proof searching.)}

>>>>

Higher Order Logic (or \emph{HOL Logic}) \cite{hollogic}, which traces
its roots back to
\emph{Edinburgh LCF} by Robin Milner and others in 1972, is a variant of
Church’s simple theory of types (STT) \cite{church1940formulation},
plus a higher order version of Hilbert's choice operator $\varepsilon$,
Axiom of Infinity, and Rank-1 (prenex) polymorphism.
HOL4 has implemented the original HOL Logic, 
while some other theorem provers in HOL family (e.g. Isabelle/HOL) have
certain extensions.
%  (they made the formal language more powerful,
% but they also bring the possibilities that the entire logic becomes
% inconsistent). 
Indeed the HOL Logic has considerable simpler logical
foundations than most other theorem provers. %, e.g. Coq. 
As a consequence,
formal theories built in HOL is easily convincible and can
also be easily ported to other proof systems,
sometimes automatically \cite{hurd2011opentheory}.

HOL4 is written in Standard ML, a single programming language which
plays three different roles:
%(The situation is quite different in other systems like Coq\footnote{\hl{The Coq Proof Assistant.} \url{https://coq.inria.fr}}.)
\begin{enumerate}
\item It serves as the underlying implementation language for the core HOL engine;\vspace{-1ex}
\item it is used to implement tactics (and tacticals) for writing proofs;\vspace{-1ex}
\item it is used as the command language of the HOL interactive shell.
\end{enumerate}
\hl{Moreover, using} the same language HOL4 users can write complex automatic
verification tools by calling HOL's theorem proving
facilities. \hl{(The formal proofs of theorems in CCS theory
are mostly done by an \emph{interactive process} closely following
their informal proofs, with minimal automatic proof searching.)}
